Shapes analysis for time series.

### 2024-05-20 NEURIPS24/appendix.tex
first results

\section{Classification datasets}
\label{appendix:classification_dataset}

All datasets were taken from UCR/UEA archives \cite{dau2019ucr,bagnall2018uea}. Among all available 
datasets\footnote{https://timeseriesclassification.com}, we selected 4 datasets related to time series 
shape comparison. All datasets were downloaded with the python package aeon\footnote{https://www.aeon-toolkit.org/en/stable/} 
which already includes the train test split. Essential dataset information is summarized in \Cref{appendix:table:datasets}.

\begin{table}[hbt!]
  \centering
  \caption{Time series datasets summary for shape based classification.}
  \begin{tabular}{lllllll}
    \toprule
    Dataset & Train size & test size & Lenghth & Number of classes & Number of dimensions & Type \\
    \midrule
    \href{https://timeseriesclassification.com/description.php?Dataset=ArrowHead}{ArrowHead} & 36 & 175 & 251 & 3 & 1 & IMAGE \\
    \href{https://timeseriesclassification.com/description.php?Dataset=ECG200}{ECG200} & 100 & 100 & 96 & 2 & 1 & ECG \\
    \href{https://timeseriesclassification.com/description.php?Dataset=GunPoint}{GunPoint} & 50 & 150 & 150 & 2 & 1 & MOTION \\
    \href{https://timeseriesclassification.com/description.php?Dataset=NATOPS}{NATOPS} & 180 & 180 & 51 & 6 & 24 & MOTION \\
    \bottomrule
  \end{tabular}
  \label{appendix:table:datasets}
\end{table}

>>>>

\section{Classification: Comparison with shape analysis methods}

In this section, we compare classification performances of TS-LDDMM with other state-of-the-art methods coming from shape analysis on 15 shape-based datasets of time-series.

\paragraph{Methods} We compare TS-LDDMM with a method from function~\cite{wu2024shape}

\paragraph{Protocole}

\begin{table}[hbt!]
  \centering
  \resizebox{\columnwidth}{!}{%
  \begin{tabular}{llrrrr}
    \toprule
     & \textbf{Dataset} & \textbf{Shape-FPCA (2024)} & \textbf{TCLR (2024)} & \textbf{LDDMM (2008)} & \textbf{TS-LDDMM (ours)} \\
    \midrule
    \multirow[c]{7}{*}{Univariate} & ArrowHead & 0.18 & 0.75 & \underline{0.84} & \textbf{0.91} \\
     & BME & 0.16 & \underline{1.00} & 0.82 & \textbf{1.00} \\
     & ECG200 & 0.40 & 0.67 & \textbf{0.81} & \underline{0.79} \\
     & FacesUCR & 0.08 & \underline{0.73} & 0.69 & \textbf{0.86} \\
     & GunPoint & 0.93 & \underline{0.97} & 0.83 & \textbf{1.00} \\
     & PhalangesOutlinesCorrect & 0.39 & \textbf{0.63} & \underline{0.53} & 0.52 \\
     & Trace & 0.55 & \underline{1.00} & 0.46 & \textbf{1.00} \\
    \cline{1-6}
    \multirow[c]{8}{*}{Multivariate} & ArticularyWordRecognition & -- & -- & \underline{0.98} & \textbf{1.00} \\
     & Cricket & -- & -- & \underline{0.77} & \textbf{0.93} \\
     & ERing & -- & -- & \underline{0.95} & \textbf{0.98} \\
     & Handwriting & -- & -- & \underline{0.22} & \textbf{0.44} \\
     & Libras & -- & -- & \underline{0.56} & \textbf{0.60} \\
     & NATOPS & -- & -- & \underline{0.82} & \textbf{0.82} \\
     & RacketSports & -- & -- & \textbf{0.83} & \underline{0.79} \\
     & UWaveGestureLibrary & -- & -- & \underline{0.72} & \textbf{0.81} \\
    \bottomrule
    \end{tabular}
    %
  }

\end{table}

\section{Classification datasets}
\label{appendix:classification_dataset}
We selected 15 shape-based datasets (7 univariates and 8 multivariates) from the from the University of East Anglia (UEA) and the University of California Riverside (UCR) Time Series Classification Repository\footnote{https://timeseriesclassification.com} \cite{dau2019ucr,bagnall2018uea}. All datasets were downloaded with the python package aeon\footnote{https://www.aeon-toolkit.org/en/stable/}. Essential datasets information are summarized in \Cref{appendix:table:datasets} and further can be found in \cite{dau2019ucr,bagnall2018uea}.

\begin{table}[hbt!]
  \centering
  \caption{UCR/UEA shape-based time series datasets for classification.}
  \resizebox{\columnwidth}{!}{%
  \begin{tabular}{lllllll}
    \toprule
    & \textbf{Dataset} &  \textbf{Size} & \textbf{Lengh} & \textbf{Number of classes} & \textbf{Number of dimensions} & \textbf{Type} \\
    \midrule
    \multirow[c]{7}{*}{Univariate} &  ArrowHead & 211 & 251 & 3 & 1 & IMAGE \\
    & BME & 180 & 128 & 3 & 1 & SIMULATED \\
    & ECG200 & 200 & 96 & 2 & 1 & ECG \\
    & FacesUCR & 2250 & 131 & 14 & 1 & IMAGE \\
    & GunPoint & 200 & 150 & 2 & 1 & MOTION \\
    & PhalangesOutlinesCorrect & 2658 & 80 & 2 & 1 & IMAGE \\
    & Trace & 200 & 275 & 4 & 1 & SENSOR \\
    \cline{1-7}
    \multirow[c]{8}{*}{Multivariate}& ArticularyWordRecognition & 575 & 144 & 25 & 9 & SENSOR \\
    & Cricket & 180 & 1197 & 12 & 6 & MOTION \\
    & ERing & 60 & 65 & 6 & 4 & SENSOR \\
    & Handwriting & 1000 & 152 & 26 & 3 & MOTION \\
    & Libras & 360 & 45 & 15 & 2 & VIDEO \\ 
    & NATOPS & 360 & 51 & 6 & 24 & MOTION \\
    & RacketSports & 303 & 30 & 4 & 6 & SENSOR \\
    & UWaveGestureLibrary & 240 & 315 & 8 &3 & SENSOR \\
    \bottomrule
  \end{tabular}
  %
  }
  \label{appendix:table:datasets}
\end{table}

### 2024-05-20 NEURIPS24/paper.tex
update
first results

\thi{
  However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g. \Cref{fig:diffeo}, thus a graph time series is more than a simple curve \cite{glaunes2008large}. Our contributions arise from this observation: we specify the class of diffeomorphisms to consider and show how to learn them. This change is fruitful in representing transformations of time series graphs as illustrated in \Cref{fig:transport}.
}

@@

\section{Notations}
We denote by integer ranges by $[k:l]=\{k,\ldots,l\}\subset \mathcal{P}(\Zset)$ and $ [l]=[1:l]$ with $k,l\in \Nset$,
by $\rmC^m(\msi,\mse)$ the set of $m$-times continously differentiable function defined on an open set $\msu$ to a normed vector space $\mse$,
 by $||u||_\infty=\inf_{x\in \msu} |u(x)| $ for any bounded function $u:\msu \to \mse$,
and by $\Nset_{>0}$ is the set of positive integers. 
%We are in a sub class of curve quotiented by temporal reparametrisation since a time series' graph can be understood as the equivalent class of t ->(t,s(t))

>>>>

% \thi{
%   However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g. \Cref{fig:diffeo}, thus a graph time series is more than a simple curve \cite{glaunes2008large}. Our contributions arise from this observation: we specify the class of diffeomorphisms to consider and show how to learn them. This change is fruitful in representing transformations of time series graphs as illustrated in \Cref{fig:transport}.
% }

@@

\thi{
  Shape analysis focuses on statistical analysis of mathematical objects invariant under some deformations like rotations, dilations, or time parameterization. The main idea is to represent these objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}. Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms. The goal is to find a well-suited Riemannian structure according to the nature of the studied object. LDDMM framework is a relevant shape analysis tool to represent curves as depicted in \cite{glaunes2008large}. However, graphs of time series are a well-structured type of curve due to the inclusion of the temporal dimension that requires specific care (\Cref{fig:diffeo}). Closely related, \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM but distinguishes from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling. Leaving the LDDMM representation, \cite{srivastava2010shape,heo2024logistic} addresses the representation of curves with the Square-Root Velocity (SRV) representation. However, the SRV representation is applied after a reparametrization of the temporal dimension on the unit length segment. Consequently, the graph structure of the time series is not respected, and the original time evolution of the time series is not encoded in the final representation. Very recently, in a functional data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by representing the original time evolution. Nevertheless, this method is made for \textit{continuous objects} and only applies to time series of \textit{same length}, making the estimation more sensitive to noise.
}
\section{Notations}
We denote by integer ranges by $[k:l]=\{k,\ldots,l\}\subset \mathcal{P}(\Zset)$ and $ [l]=[1:l]$ with $k,l\in \Nset$,
by $\rmC^m(\msi,\mse)$ the set of $m$-times continously differentiable function defined on an open set $\msu$ to a normed vector space $\mse$,
 by $||u||_\infty=\inf_{x\in \msu} |u(x)| $ for any bounded function $u:\msu \to \mse$,
and by $\Nset_{>0}$ is the set of positive integers. 
%We are in a sub class of curve quotiented by temporal reparametrisation since a time series' graph can be understood as the equivalent class of t ->(t,s(t))
